\documentclass[12pt, psamsfonts]{amsart}

%-------Packages---------
\usepackage{amssymb,amsfonts}
\usepackage{fullpage}
\usepackage{todonotes}
\usepackage{physics}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{mathrsfs}
\usepackage{theoremref}
\usepackage{graphicx}
\usepackage[bookmarks]{hyperref}

%--------Theorem Environments--------
%theoremstyle{plain} --- default
\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{conj}[thm]{Conjecture}
\newtheorem{quest}[thm]{Question}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{defns}[thm]{Definitions}
\newtheorem{con}[thm]{Construction}
\newtheorem{exmp}[thm]{Example}
\newtheorem{exmps}[thm]{Examples}
\newtheorem{notn}[thm]{Notation}
\newtheorem{notns}[thm]{Notations}
\newtheorem{addm}[thm]{Addendum}
\newtheorem*{exer}{Exercise}

\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}
\newtheorem{rems}[thm]{Remarks}
\newtheorem{warn}[thm]{Warning}
\newtheorem{sch}[thm]{Scholium}

\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Id}{Id}

\makeatletter
\let\c@equation\c@thm
\makeatother
\numberwithin{equation}{section}

\bibliographystyle{plain}

\begin{document}

\title{Math 601 (Due 10/2)}
\author{Hidenori Shinohara}
\maketitle

\tableofcontents

\section{Rings of Fractions}
\begin{exer}{(Problem 1 (iii))}
  Prove that the natural map $i: R \rightarrow S^{-1}R$, which maps $r$ to $\frac{r}{1}$ is an injective ring homomorphism.
\end{exer}

\begin{proof}
  $ $
  \begin{itemize}
    \item
      Ring homomorphism?
      \begin{itemize}
        \item
          For all $r, s \in R$, $i(rs) = \frac{rs}{1} = \frac{r}{1} \cdot \frac{s}{1} = i(r)i(s)$.
        \item
          For all $r, s \in R$, $i(r + s) = \frac{r + s}{1} = \frac{r}{1} + \frac{s}{1} = i(r) + i(s)$.
      \end{itemize}
      Therefore, $i$ is indeed a ring homomorphism.
    \item
      Injective?
      It suffices to check that $\ker(i) = \{ 1 \}$.
      Let $r \in R$ such that $\ker(r)$ is the multiplicative identity in $S^{-1}R$.
      By definition, $\ker(r) = \frac{1}{1}$.
      Thus $\frac{r}{1} = \frac{1}{1}$, so $r \cdot 1 - 1 \cdot 1 = 0$.
      This means $r = 1$, so $\ker(i) = \{ 1 \}$.
  \end{itemize}
  Therefore, $i$ is indeed an injective ring homomorphism.
\end{proof}

\begin{exer}{(Problem 1(iv))}
  Prove that given a ring homomorphism $h: R \rightarrow T$, such that $h(s) \in T^{\star}$ for every $s \in S$, there exists a unique ring homomorphism $\lambda: S^{-1}R \rightarrow T$, such that $h = \lambda \circ i$.
\end{exer}

\begin{proof}
  Suppose such a $\lambda$ exists.
  Then for all $r \in R$, $h(r) = (\lambda \circ i)(r) = \lambda(r / 1)$.
  Therefore, $\lambda(r / 1) = h(r)$.
  Let $s \in S$.
  Then $1_T = \lambda(1 / 1) = \lambda((s / 1) \cdot (1 / s)) = \lambda(s / 1)\lambda(1 / s)$.
  Therefore, $\lambda(1 / s) = \lambda(s / 1)^{-1} = h(s)^{-1}$.
  This implies that $\lambda(r / s) = \lambda(r / 1)\lambda(1 / s) = h(r)h(s)^{-1}$.

  In other words, if such a $\lambda$ exists, it must map $r / s$ to $h(r)h(s)^{-1}$.
  This proves the uniqueness.
  We will show that such a function is indeed well defined and it is a ring homomorphism.
  \begin{itemize}
    \item
      Well-defined?
      Since $h(s) \in T^{\star}$ for each $s \in S$, $h(s)^{-1}$ is well defined.
      Let $r / s = r' / s' \in S^{-1}R$ be given.
      Then $rs' = r's$.
      Since $h$ is a ring homomorphism, $h(r)h(s') = h(r')h(s)$.
      Therefore, $\lambda(r / s) = h(r)h(s)^{-1} = h(r')h(s')^{-1} = \lambda(r' / s')$.
    \item
      Ring homomorphism?
      Let $r / s, r' / s' \in S^{-1}R$.
      \begin{align*}
        \lambda(\frac{r}{s} \cdot \frac{r'}{s'})
          &= \lambda(\frac{rr'}{ss'}) \\ 
          &= h(rr')h(ss')^{-1} \\
          &= h(r)h(r')h(s)^{-1}h(s')^{-1} \\
          &= h(r)h(s)^{-1}h(r')h(s')^{-1} \\
          &= \lambda(\frac{r}{s})\lambda(\frac{r'}{s'}). \\
        \lambda(\frac{r}{s} + \frac{r'}{s'})
          &= \lambda(\frac{rs' + r's}{ss'}) \\
          &= h(rs' + r's)h(ss')^{-1} \\
          &= (h(r)h(s') + h(r')h(s))h(s)^{-1}h(s')^{-1} \\
          &= h(r)h(s)^{-1} + h(r')h(s')^{-1} \\
          &= \lambda(\frac{r}{s}) + \lambda(\frac{r'}{s'}).
      \end{align*}
    \item
      Commutes?
      For any $r \in R$, $\lambda(i(r)) = \lambda(r / 1) = h(r)h(1)^{-1} = h(r)$.
      Therefore, $\lambda \circ i$ is indeed $h$.
  \end{itemize}
\end{proof}

\section{The Quadratic Equation $x^2 - 2y^2 = n$}
\begin{exer}{(Problem 15)}
  Find a solution to $x^2 - 2y^2 = 7$.
\end{exer}

\begin{proof}
  $3^2 - 2 \cdot 1^2 = 9 - 2 = 7$.
  Thus $(x, y) = (3, 1)$ is a solution to $x^2 - 2y^2 = 7$.
\end{proof}

\begin{exer}{(Problem 16)}
  Is 7 irreducible in $\mathbb{Z}[\sqrt{2}]$?
  If not, find a factorization into irreducible elements.
\end{exer}

\begin{proof}
  By Problem 3 from the previous assignment, we know that $\alpha \in \mathbb{Z}[\sqrt{2}]$ is a unit if and only if $N(\alpha) = \pm 1$.
  We will use this result in this solution.

  By Problem 15, we know that $7 = (3 + \sqrt{2})(3 - \sqrt{2})$.
  Since $N(3 + \sqrt{2}) = N(3 - \sqrt{2}) = 7 \ne \pm 1$, 7 can be expressed as a product of two non-unit elements, so 7 is not irreducible.

  Suppose $3 + \sqrt{2} = (a + b\sqrt{2})(c + d\sqrt{2})$ for some $a, b, c, d \in \mathbb{Z}$.
  By Problem 2 from the previous assignment, we know that $N(3 + \sqrt{2}) = N(a + b\sqrt{2})N(c + d\sqrt{2})$.
  Since $N$ maps $\mathbb{Z}[\sqrt{2}]$ into integers, exactly one of $N(a + b\sqrt{2})$ and $N(c + d\sqrt{2}$ must be 1 or -1, and the other one is 7 or -7.
  Therefore, one of $a + b\sqrt{2}$ or $c + d\sqrt{2}$ is a unit, so $3 + \sqrt{2}$ is irreducible.

  Similarly, if $3 - \sqrt{2} = (a' + b'\sqrt{2})(c' + d'\sqrt{2})$, then $7 = N(3 - \sqrt{2}) = N(a' + b'\sqrt{2})N(c' + d'\sqrt{2})$.
  Therefore, one of $a' + b'\sqrt{2}$ or $c' + d'\sqrt{2}$ is a unit, so $3 - \sqrt{2}$ is irreducible.
\end{proof}

\begin{exer}{(Problem 17)}
  Let $p \in \mathbb{Z} \setminus \{ 0 \}$ and suppose $\alpha\beta = p$ in $\mathbb{Z}[\sqrt{2}]$.
  Show that $\beta = c\gamma(\alpha)$ with $c \in \mathbb{Q}$.
\end{exer}

\begin{proof}
  Choose $a, b, c, d \in \mathbb{Z}$ such that $a + b\sqrt{2} = \beta, c + d\sqrt{2} = \alpha$.
  Since $\alpha\beta = p \ne 0$, $\alpha \ne 0$.
  This implies at least one of $c$ or $d$ is nonzero.
  Therefore, $\gamma(\alpha) = c - d\sqrt{2} \ne 0$.

  We have $\alpha\beta = (ac + 2bd) + \sqrt{2}(ad + bc)$.
  Since $\alpha\beta \in \mathbb{Z}$, $ad + bc = 0$.

  \begin{align*}
    \frac{\beta}{\gamma(\alpha)}
      &= \frac{a + b\sqrt{2}}{c - d\sqrt{2}} \\
      &= \frac{(a + b\sqrt{2})(c + d\sqrt{2})}{c^2 - 2d^2} \\
      &= \frac{(ac + 2bd) + (ad + bc)\sqrt{2}}{c^2 - 2d^2} \\
      &= \frac{ac + 2bd}{c^2 - 2d^2}.
  \end{align*}

  Therefore, $\frac{\beta}{\gamma(\alpha)}  = \frac{ac + 2bd}{c^2 - 2d^2} \in \mathbb{Q}$.
  In other words, $\beta = \frac{ac + 2bd}{c^2 - 2d^2}\gamma(\alpha)$.
\end{proof}

\begin{exer}{(Problem 18)}
  Let $p \in \mathbb{Z}$ be an odd prime.
  Show that $p = N(\alpha)$ for some $\alpha \in \mathbb{Z}[\sqrt{2}]$ if and only if $p$ is not irreducible as an element of $\mathbb{Z}[\sqrt{2}]$.
\end{exer}

\begin{proof}
  By Problem 3 from the previous assignment, we know that $\alpha \in \mathbb{Z}[\sqrt{2}]$ is a unit if and only if $N(\alpha) = \pm 1$.
  We will use this result in this solution.

  Suppose $p = N(\alpha)$ for some $\alpha \in \mathbb{Z}[\sqrt{2}]$.
  Since $N(\alpha) = \alpha\gamma(\alpha)$, $p$ can be written as a product of $\alpha$ and $\gamma(\alpha)$.
  \begin{itemize}
    \item
      $N(\alpha) = p \ne \pm 1$, so $\alpha$ is not a unit.
    \item
      Since $N(\gamma(\alpha)) = \gamma(\alpha)\gamma(\gamma(\alpha)) = \gamma(\alpha)\alpha = N(\alpha) = p \ne \pm 1$, $\gamma(\alpha)$ is not a unit.
  \end{itemize}
  Therefore, $p$ is a product of two non-unit elements $\alpha, \gamma(\alpha)$, so $p$ is not irreducible.

  On the other hand, suppose that $p$ is not irreducible as an element of $\mathbb{Z}[\sqrt{2}]$.
  Then $p = \alpha\beta$ where $\alpha, \beta \in \mathbb{Z}[\sqrt{2}]$ are non-unit elements.
  Then $N(p) = N(\alpha)N(\beta)$.

  \begin{itemize}
    \item
      $N(p) = p^2$ because $p$ is an integer.
    \item
      $N(\alpha) \ne \pm 1$ because $\alpha$ is not a unit.
    \item
      $N(\beta) \ne \pm 1$ because $\beta$ is not a unit.
  \end{itemize}
  Since $N(\alpha), N(\beta)$ are both integers, $N(\alpha) = N(\beta) = p$ or $N(\alpha) = N(\beta) = -p$.
  If $N(\alpha) = p$, then we are done.
  If $N(\alpha) = -p$, then $N(\alpha(1 + \sqrt{2})) = N(\alpha)N(1 + \sqrt{2}) = (-p)(-1) = p$.
\end{proof}

\begin{exer}{(Problem 19)}
  Let $p \in \mathbb{Z}$ be an odd prime.
  Show that $x^2 - 2y^2 = p$ has a solution if and only if $p$ is not irreducible in $\mathbb{Z}[\sqrt{2}]$.
\end{exer}

\begin{proof}
  Let an odd prime $p$ be given.
  There exists an $\alpha \in \mathbb{Z}[\sqrt{2}]$ such that $p = N(\alpha)$ if and only if there exist $x, y \in \mathbb{Z}$ such that $p = x^2 - 2y^2$ because $N(x + \sqrt{2}y) = x^2 - 2y^2$.
  By combining this with the results of Problem 18, we have $x^2 - 2y^2 = p$ has a solution if and only if $p$ is not irreducible in $\mathbb{Z}[\sqrt{2}]$.
\end{proof}

\end{document}
