\documentclass[12pt, psamsfonts]{amsart}

%-------Packages---------
\usepackage{listings}
\usepackage{breqn}
\usepackage{amssymb,amsfonts}
\usepackage{fullpage}
\usepackage{tikz-cd}
\usepackage{todonotes}
\usepackage{physics}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{mathrsfs}
\usepackage{theoremref}
\usepackage{graphicx}
\usepackage[bookmarks]{hyperref}

%--------Theorem Environments--------
%theoremstyle{plain} --- default
\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{conj}[thm]{Conjecture}
\newtheorem{quest}[thm]{Question}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{defns}[thm]{Definitions}
\newtheorem{con}[thm]{Construction}
\newtheorem{exmp}[thm]{Example}
\newtheorem{exmps}[thm]{Examples}
\newtheorem{notn}[thm]{Notation}
\newtheorem{notns}[thm]{Notations}
\newtheorem{addm}[thm]{Addendum}
\newtheorem*{exer}{Exercise}

\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}
\newtheorem{rems}[thm]{Remarks}
\newtheorem{warn}[thm]{Warning}
\newtheorem{sch}[thm]{Scholium}

\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Id}{Id}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\ord}{ord}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\Gal}{Gal}

\makeatletter
\let\c@equation\c@thm
\makeatother
\numberwithin{equation}{section}

\bibliographystyle{plain}

\begin{document}

\title{Math 601 (Due 11/22)}
\author{Hidenori Shinohara}
\maketitle

\tableofcontents
\section{THE THEOREM ON SYMMETRIC POLYNOMIALS}

\begin{exer}{(Problem 1)}
  By substituting $u_4 = 0$, we get $u_1^2u_2u_3 + u_1u_2^2u_3 + u_1u_2u_3^2 = s_3s_1$.
  $s_3s_1$ with 4 variables expands to $u_{1}^{2} u_{2} u_{3} + u_{1}^{2} u_{2} u_{4} + u_{1}^{2} u_{3} u_{4} + u_{1} u_{2}^{2} u_{3} + u_{1} u_{2}^{2} u_{4} + u_{1} u_{2} u_{3}^{2} + 4 u_{1} u_{2} u_{3} u_{4} + u_{1} u_{2} u_{4}^{2} + u_{1} u_{3}^{2} u_{4} + u_{1} u_{3} u_{4}^{2} + u_{2}^{2} u_{3} u_{4} + u_{2} u_{3}^{2} u_{4} + u_{2} u_{3} u_{4}^{2}$.
  Then $s_3s_1 - f$ where $f$ is the original polynomial gives us $4u_1u_2u_3u_4 = 4s_4$.
  Therefore, $f = s_3s_1 - 4s_4$.
\end{exer}

\begin{exer}{(Problem 2)}
  We are given that $\abs{M - xI} = x^3 - ax^2 + bx - c$.
  This implies that $\abs{M - (-x)I} = -x^3 - ax^2 - bx - c$.
  Since the determinant function preserves multiplication, $\abs{M - xI}\abs{M - (-x)I} = \abs{M^2 - x^2I}$.
  This implies $\abs{M^2 - x^2I} = -x^6 + (a^2 - 2b)x^4 + (-b^2 + 2ac)x^2 + c^2$.
  Therefore, the characteristic polynomial of $M$ is $-x^3 + (a^2 - 2b)x^2 + (-b^2 + 2ac)x + c^2$.
  \todo[inline,caption={}]{
    $\abs{xI - M}$ is what I need to calculate.
  }
\end{exer}

\section{Galois Theory VI}

\begin{exer}{(Problem 3)}
  \begin{enumerate}[label=(\alph*)]
    \item
      $\{ (123), (132), e \}$ is clearly a subgroup of the stabilizer group $S_v$ of $v$.
      Since $(12) \notin S_v$, $3 \leq \abs{S_v} \leq 5$.
      By Lagrange's Theorem, $S_v = \ev{(123)}$. 
    \item
      By (i), $S_3v$ contains only $[S_3:S_v] = 2$ elements.
      Thus $v' = (12) \cdot v = u_2u_1^2 + u_1u_3^2 + u_3u_2^2$.
    \item
      By substituting $u_3 = 0$ for $v + v'$, we get $u_1u_2^2 + u_2u_1^2 = s_1s_2$.
      Then $v + v' - s_1s_2 = -3u_1u_2u_3 = -3s_3$.
      Therefore, $v + v' = s_1s_2 - 3s_3$.
    \item
      We will use the fundamental theorem of Galois Theory.
      $F(v) = K^{\ev{(123)}}$, so $\abs{\ev{(123)}} = 3 = [K:F(v)]$.
      Moreover, $\abs{\ev{\Gal(K/F)}} = [K:F]$.
      Therefore, $[F(v):F] = [K:F]/[K:F(v)] = \abs{\Gal(K/F)}/3$.
    \item
      Calculation shows that $vv' = 9s_3^2 + s_3s_1^3 - 6s_3s_1s_2 + s_2^3$.
      By substituting $s_1 = 0, s_2 = p, s_3 = q$, we get $9q^2 + p^3$.
    \item
      Since $A_3$ is the only proper transitive subgroup of $S_3$, $\Gal(K/F) = S_3$ if and only if $\sigma \in \Gal(K/F)$ where $\sigma$ corresponds to the permutation $(12)$.
      (i.e., $u_1 \mapsto u_2, u_2 \mapsto u_1$.)
      $v, v'$ are not fixed by $\sigma$, so $v, v' \notin F$ if $\Gal(K/F) = S_3$.
      $v, v'$ are fixed by every permutation if $\Gal(K/F) = A_3$ because it is generated by $\sigma'$ that corresponds to $(123)$.
      Therefore, we can conclude that $\Gal(K/F) \ne S_3$ if and only if $v, v' \in F$.
      $v, v' \in F$ if and only if $(y - v)(y - v')$ factors in $F$.
      Therefore, $h(y) = y^2 - (v + v')y + vv' = y^2 + 3qy + (9q^2 + p^3)$ is the desired polynomial.
    \item
      The discriminant is $(3q)^2 - 4(9q^2 + p^3) = 9q^2 - 36q^2 - 4p^3 = -27q^2 - 4p^3$.
  \end{enumerate}
\end{exer}

\begin{exer}{(Problem 4)}
  \begin{enumerate}[label=(\alph*)]
    \item 
      The discriminant can be expressed as $-4s_{1}^{3}s_{3}+s_{1}^{2}s_{2}^{2}+18s_{1}s_{2}s_{3}-4s_{2}^{3}-27s_{3}^{2}$.
      By substituting $s_1 = 1, s_2 = -2, s_3 = -1$, we get 49.
      \lstinputlisting[language=python]{galois_problem4.py}
      Since the discriminant is a square, the Galois group is isomorphic to $A_3$.

  \end{enumerate}
\end{exer}

\begin{exer}{(Problem 5)}
  \begin{enumerate}[label=(\alph*)]
    \item 
      \todo[inline,caption={}]{
      }
    \item
      $x^4 + x + 1$ is irreducible because
      \begin{itemize}
        \item
          It does not have a linear factor by the rational root theorem.
        \item
          If it factors into two rational quadratic polynomials, they will factor into two monic integer quadratic polynomials, namely, $x^2 + ax + b$ and $x^2 - ax + 1/b$ based on the coefficients.
          This implies $b = \pm 1$.
          Since the coefficient of x is 1, $-ab + a/b = 1$, but this implies $b \ne \pm 1$.
      \end{itemize}
      We will use the discussion presented in the Galois Theory IV handout.
      By (i), the discriminant is 229, so $h(y) = y^2 - 229$.
      Also, $g(y) = y^3 - 4y - 1$ since $a = b = 0, c = -1, d = 1$.
      Therefore, both $h(y)$ and $g(y)$ are irreducible, so the Galois group is $S_4$.
    \item
      It does not have a linear factor by the rational root theorem.
      Based on coefficients, if it factors into quadratic polynomials, it will be $(x^2 + ax + b)(x^2 - ax + c)$ for some $a, b, c \in \mathbb{Z}$ by Gauss' lemma.
      This gives $bc = 12$ and $-ab + ac = -8$, so $a(c - 12/c) = -8$.
      This is a quadratic polynomial in $c$ with the discriminant $64 - 48a$.
      This must be a square for $c$ to exist.
      By checking each possible value of $a$, we get $64 - 48 \cdot -8 = 448, 64 - 48 \cdot -4 = 256, 64 - 48 \cdot -2 = 160, 64 - 48 \cdot -1 = 112, 64 - 48 \cdot 1 = 16$.
      (For other $a$, $64 - 48a < 0$.)
      Thus the only two possible values are $a = 1, -4$.
      $a = 1$ gives $c - b = -8$ and $bc = 12$, which we can confirm to be impossible by examining the divisors of 12.
      Similarly, $a = -4$ gives $c - b = 2$ and $bc = 12$ and this is impossible to satisfy.
      Therefore, $x^4 - 8x + 12$ is irreducible over $\mathbb{Q}$.
    \item
      Again, we will use the discussion presented in the Galois Theory IV handout.
      By calculating the discriminant, we have $h(y) = h(y) = y^{2} - 331776$ and $g(y) = y^{3} - 48 y - 64$.
      $h(y)$ factors as $576^2 = 331776$.
      $g(y)$ does not factor by the rational root theorem.
      Therefore, the Galois group is $A_4$.
  \end{enumerate}
\end{exer}

\section{Galois Theory V(Further exercises)}
\section{Galois Theory V}

\begin{exer}{(Problem 1)}
  \begin{enumerate}[label=(\alph*)]
    \item 
      $D_4$ is a subgroup of $S_4$ and the Galois Theory V handout states that $S_4$ is solvable and any subgroup of a finite solvable group is solvable.
    \item
      If $S_5$ is solvable, $A_5 \leq S_5$ is solvable.
      However, as stated in the handout, $A_5$ is not solvable.
  \end{enumerate}
\end{exer}

\begin{exer}{(Problem 2)}
  \begin{enumerate}[label=(\alph*)]
    \item 
      \todo[inline,caption={}]{
      }
    \item
      $G$ has a subgroup $G_1 = \{(1),$ $(12)(35),$ $(12345),$ $(13)(45),$ $(13524),$ $(14)(23),$ $(14253),$ $(15)(24),$ $(15432),$ $(25)(34)\}$.
      of index 2.
      $G_1$ has a subgroup $G_2 = \{(1),$ $(12345),$ $(13524),$ $(14253),$ $(15432)\}$ of index 2.
      $G_2$ is abelian, so we can pick $G_3 = \{(1)\}$.
      Then $G_3 \subset G_2 \subset G_1 \subset G$ is a filteration. 

  \end{enumerate}
\end{exer}

\end{document}


