\documentclass[12pt, psamsfonts]{amsart}

%-------Packages---------
\usepackage{amssymb,amsfonts}
\usepackage{semantic}
\usepackage{fullpage}
\usepackage{tikz-cd}
\usepackage{todonotes}
\usepackage{physics}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{mathrsfs}
\usepackage{theoremref}
\usepackage{graphicx}
\usepackage[bookmarks]{hyperref}

%--------Theorem Environments--------
%theoremstyle{plain} --- default
\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{conj}[thm]{Conjecture}
\newtheorem{quest}[thm]{Question}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{defns}[thm]{Definitions}
\newtheorem{con}[thm]{Construction}
\newtheorem{exmp}[thm]{Example}
\newtheorem{exmps}[thm]{Examples}
\newtheorem{notn}[thm]{Notation}
\newtheorem{notns}[thm]{Notations}
\newtheorem{addm}[thm]{Addendum}
\newtheorem*{exer}{Exercise}

\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}
\newtheorem{rems}[thm]{Remarks}
\newtheorem{warn}[thm]{Warning}
\newtheorem{sch}[thm]{Scholium}

\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Id}{Id}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\ord}{ord}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\Gal}{Gal}
\DeclareMathOperator{\Int}{int}

\makeatletter
\let\c@equation\c@thm
\makeatother
\numberwithin{equation}{section}

\bibliographystyle{plain}

\begin{document}

\title{Math 633 Homework 3}
\author{Hidenori Shinohara}
\maketitle

\begin{exer}{(Problem 1)}
  A simply connected space is clearly piecewise smooth simply connected.
  Let $\Omega$ be piecewise smooth simply connected and $\gamma_1, \gamma_2: [0, 1] \rightarrow \Omega$ be two continuous curves with the same end points.
  Since $\Omega$ is open, $\gamma_1(t)$ has an open ball around it that is contained in $\Omega$ for each $t \in [0, 1]$.
  Since $[0, 1]$ is compact and $\gamma_1$ is continuous, $\gamma_1([0, 1])$ is compact.
  Hence, there is a finite partition $0 = t_0 < t_1 < \cdots < t_{n - 1} < t_n = 1$ such that $\gamma_1([t_i, t_{i + 1}])$ is contained in an open ball $\subset \Omega$ for each $i$.
  Then $\gamma_1$ is homotopic to the curve $\gamma_{1'}$ that consists of $n$ straight lines, $i$th of which is the line between $\gamma_1(t_i)$ and $\gamma_1(t_{i - 1})$ where $i = 1, \cdots, n$.
  This can be shown by the ``straight-line" homotopy because $\gamma_1([t_{i - 1}, t_{i}])$ and the $i$th straight line are in an open ball contained in $\Omega$.

  A similar argument can be applied to show that $\gamma_2$ is homotopic to a curve $\gamma_{2'}$ that consists of finitely many straight lines.
  A curve consisting of finitely many straight lines is clearly piecewise smooth.

  Therefore, $\gamma_1 \sim \gamma_{1'} \sim \gamma_{2'} \sim \gamma_2$.
  Thus $\Omega$ is simply connected.
\end{exer}

\begin{exer}{(Problem 2)}
  Define $T(x, y) = x + iy$.
  \begin{align*}
    \int_{S} fdz
      &= \int_{0}^{1} f(t)dt + \int_{0}^{1} f(it)(it)'dt + \int_{0}^{1} f(1 + it)(1 + it)'dt + \int_{0}^{1} f(t + i)(t + i)'dt \\
      &= \int_{0}^{1} f(t) + f(t + i)dt + i\int_{0}^{1} f(it) + f(1 + it) dt \\
      &= \int_{0}^{1} f(T(x, 0)) + f(T(x, 1)) dx + i\int_{0}^{1} f(T(0, y)) + f(T(1, y)) dy \\
      &= \int_{0}^{1} u(T(x, 0)) + u(T(x, 1)) dx + i\int_{0}^{1} u(T(0, y)) + u(T(1, y)) dy  \\
      &\nbsp + i\int_{0}^{1} v(T(x, 0)) + v(T(x, 1)) dx - \int_{0}^{1} v(T(0, y)) + v(T(1, y)) dy \\
      &= \int_{0}^{1} u(T(x, 0)) + u(T(x, 1)) dx - \int_{0}^{1} v(T(0, y)) + v(T(1, y)) dy \\
      &\nbsp + i(\int_{0}^{1} u(T(0, y)) + u(T(1, y)) dy + \int_{0}^{1} v(T(x, 0)) + v(T(x, 1)) dx) \\
      &= \int_{S} u \circ T dx + \int_{S} -v \circ T dy + i(\int_{S} v \circ T dx + \int_{S} u \circ T) \\
      &= \int_{\Int S} -\frac{\partial v}{\partial x} - \frac{\partial u}{\partial y} + i(\int_{\Int S} \frac{\partial u}{\partial x} - \frac{\partial v}{\partial y}) \\
      &= 0.
  \end{align*}

\end{exer}

\begin{exer}{(Problem 3)}
  Let $\gamma_1, \gamma_2, \gamma_3$ denote the real part of the contour, the arc, and the rest, respectively.
  $\int_{\gamma_1 + \gamma_2 + \gamma_3} e^{z^2} = 0$ because $e^{z^2}$ is entire.

  Let $C = \int_{0}^{\infty} \cos(x^2) dx, S = \int_{0}^{\infty} \sin(x^2) dx$.

  \begin{itemize}
    \item
      By setting $\gamma_1(t) = t$ where $t \in [0, R]$, we obtain $\lim_{R \rightarrow \infty} \int_{0}^{R} e^{-t^2} dt = \sqrt{\pi} / 2$.
      Thus the integration over $\gamma_1$ is $\sqrt{\pi} / 2$.
    \item
      Let $\gamma_2(t) = Re^{it}$ with $t \in [0, \pi / 4]$.
      Then $\abs{\int_0^{\pi/4} e^{-\gamma_2^2} \gamma_2' dt} \leq \int_0^{\pi/4} \abs{e^{-\gamma_2^2} \gamma_2'} dt = R\int_{0}^{\pi/4} \abs{e^{-R^2e^{2it}}} dt$.
      Since $\abs{e^z} = e^{\Re(z)}$, it suffices to calculate $R\int_{0}^{\pi / 4} e^{-R^2\cos(2t)} dt$.
      $\cos(2t) \geq 1 - \frac{4t}{\pi} \geq 0$ on $[0, \pi / 4]$, where the $1 - \frac{4t}{\pi}$ denotes the straight-line approximation of $\cos(2t)$ over the given interval.
      Thus the integral is bounded by $R\int_{0}^{\pi / 4}e^{-R^2(1-\frac{4t}{\pi})} dt$.
      \begin{align*}
        R\int_{0}^{\pi / 4}e^{-R^2(1-\frac{4t}{\pi})} dt
          &= \frac{R}{e^{R^2}} (\frac{\pi}{4R^2}e^{\frac{4R^2t}{\pi}} \Big\vert^{\pi/4}_0) \\
          &= \frac{\pi}{4Re^{R^2}} (e^{R^2} - 1) \\
          &= \frac{\pi}{4R} (1 - \frac{1}{e^{R^2}}).
      \end{align*}

      When $R \rightarrow 0$, it is clear that the limit approaches 0.
      Therefore, $\int_{\gamma} e^{-z^2} dz = 0$.
    \item
      Let $\gamma_3(t) = te^{i\pi / 4}$ with $0 \leq t \leq R$.
      To simplify the calculation , $\gamma_3$ is oriented in the opposite way, so we will multiply $-1$ in the end.
      \begin{align*}
        \lim_{R \rightarrow \infty} \int_0^R e^{-\gamma_3^2(t)}\gamma_3'(t) dt
          &= \frac{1 + i}{\sqrt{2}} \lim_{R \rightarrow \infty} \int_0^R e^{-t^2e^{\pi i / 2}} dt \\
          &= \frac{1 + i}{\sqrt{2}} \lim_{R \rightarrow \infty} \int_0^R e^{-t^2i} dt \\
          &= \frac{1 + i}{\sqrt{2}} \lim_{R \rightarrow \infty} \int_0^R \cos(-t^2) + i\sin(-t^2) dt \\
          &= \frac{1 + i}{\sqrt{2}} [\lim_{R \rightarrow \infty} \int_0^R \cos(-t^2) + \lim_{R \rightarrow \infty} \int_0^R i\sin(-t^2) dt] \\
          &= \frac{1 + i}{\sqrt{2}} [\lim_{R \rightarrow \infty} \int_0^R \cos(t^2) - \lim_{R \rightarrow \infty} \int_0^R i\sin(t^2) dt] \\
          &= \frac{1 + i}{\sqrt{2}} [C - iS].
      \end{align*}
  \end{itemize}
  By putting these together and comparing the real and imaginary parts, we obtain a system of equations:
  \begin{align*}
    0 &= \frac{\pi}{2} - \frac{C}{\sqrt{2}} - \frac{S}{\sqrt{2}} \\
    0 &= -C + S.
  \end{align*}

  By solving it, we obtain $S = C = \sqrt{2\pi} / 4$.
\end{exer}

\begin{exer}{(Problem 4)}
  \begin{itemize}
    \item
      $\Omega_1$ is simply connected because any two continuous curves with the same end points are joined by the straight-line homotopy.
    \item
      $\Omega_2$ is not simply connected because $\Omega_2$ is homeomorphic to $S^1$ which has a nontrivial fundamental group.
      In other words, $\phi: \theta \mapsto (a + b)e^{2\pi i\theta}/2$ is a continuous curve in $\Omega$ that is not homotopic to the constant curve at $(a + b) / 2$.
    \item
      $\Omega_3$ is not simply connected because it is not connected.
  \end{itemize}
\end{exer}

\end{document}


