\documentclass[12pt, psamsfonts]{amsart}

%-------Packages---------
\usepackage{amssymb,amsfonts}
\usepackage{fullpage}
\usepackage{tikz-cd}
\usepackage{todonotes}
\usepackage{physics}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{mathrsfs}
\usepackage{theoremref}
\usepackage{graphicx}
\usepackage[bookmarks]{hyperref}

%--------Theorem Environments--------
%theoremstyle{plain} --- default
\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{conj}[thm]{Conjecture}
\newtheorem{quest}[thm]{Question}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{defns}[thm]{Definitions}
\newtheorem{con}[thm]{Construction}
\newtheorem{exmp}[thm]{Example}
\newtheorem{exmps}[thm]{Examples}
\newtheorem{notn}[thm]{Notation}
\newtheorem{notns}[thm]{Notations}
\newtheorem{addm}[thm]{Addendum}
\newtheorem*{exer}{Exercise}

\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}
\newtheorem{rems}[thm]{Remarks}
\newtheorem{warn}[thm]{Warning}
\newtheorem{sch}[thm]{Scholium}

\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Id}{Id}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\ord}{ord}

\makeatletter
\let\c@equation\c@thm
\makeatother
\numberwithin{equation}{section}

\bibliographystyle{plain}

\begin{document}

\title{Algebraic Topology Examples 1}
\author{Hidenori Shinohara}
\maketitle

\section{Covering spaces}

\begin{exmps}
  $S^1$ can be seen as a 2-sheeted covering space of $S^1$.
  \begin{itemize}
    \item
      Define a 2-sheeted covering map $p: S^1 \rightarrow S^1$.
    \item
      Let $X = S^1, \tilde{X} = S^1$.
      Compute $\pi_1(X), \pi_1(\tilde{X}), p_*(\pi_1(\tilde{X}))$.
    \item
      Compute the deck transformation group of $\tilde{X}$.
    \item
      Is $\tilde{X}$ normal?
    \item
      If $\tilde{X}$ is normal, check if $G(\tilde{X})$ abelian.
    \item
      Confirm that the groups calculated above are indeed correct by using Proposition 1.39(b). 
    \item
      Is there a group $G$ such that $p$ can be expressed as $S^1 \mapsto S^1 / G$?
      Is the action properly discontinuous?
  \end{itemize}
\end{exmps}

\begin{exmps}
  $S^1$ can be seen as 2-sheeted, 3-sheeted covering spaces of $S^1$.
  Determine if they are isomorphic to each other.
\end{exmps}

\begin{exmps}
  Given a covering map $p: (\tilde{X}, \tilde{x_0}) \rightarrow (X, x_0)$ with $\tilde{X}$ path connected, any deck transformation $\tau$ is uniquely determined by $\tau(\tilde{x_0})$.
  (This is from 10/2)

  Check that this is indeed true by looking at the universal covering of $S^1 \vee S^1$.

  Find $(X, x_0), (\tilde{X}, \tilde{x_0})$ such that
  \begin{itemize}
    \item
      $\tilde{X}$ is a covering space of $X$.
    \item
      $\tilde{X}$ is not path connected.
    \item
      A deck transformation $\tau$ cannot be uniquely determined by $\tau(\tilde{x_0})$.
  \end{itemize}
\end{exmps}


\begin{exmps}
  Consider the following covering spaces:
  
  \begin{itemize}
    \item
      $p: S^1 \vee S^1 \vee S^1 \vee S^1 \vee S^1 \rightarrow S^1 \vee S^1$.
    \item
      $p: S^1 \vee S^1 \vee S^1 \vee S^1 \rightarrow S^1 \vee S^1$.
    \item
      $p: S^1 \vee S^1 \vee S^1 \rightarrow S^1 \vee S^1$.
    \item
      $p: S^1 \vee S^1 \rightarrow S^1 \vee S^1$.
  \end{itemize}

  For each of the above,

  \begin{itemize}
    \item
      Determine if it is normal by using the definition, and come up with a proof by picture.
    \item
      Calculate the deck transformation group.
    \item
      Calculate $H = p_*(\pi_1(\tilde{X})) \subset \pi_1(X)$.
      Is there any connection between the deck transformation group and $H$?
    \item
      Apply Proposition 1.39(a) to confirm that we get the same answer.
  \end{itemize}
\end{exmps}

\end{document}
